---
---
title: "Supply Chain Vector Analysis"
author: "Guibril Ramde"
format: pdf
---

## Tasks

## Part 1 Inventory

```{r}
library(tidyverse)
#Inventory data load
data = ("https://raw.githubusercontent.com/japhet125/Data_Science_605_Hws/refs/heads/main/inventory.csv")
inventory_data = read.csv(data)
inventory_data
```

## Part 2 Demand

```{r}
#Demand data load
data2 = ("https://raw.githubusercontent.com/japhet125/Data_Science_605_Hws/refs/heads/main/demand.csv")
demand_data2 = read.csv(data2)
demand_data2
```

# Tasks

# Part A: Inventory Analysis

## Using vector operations:

```{r}
inventory_data$Region1
inventory_data$Region2
inventory_data$Region3
```

# Calculate total inventory available in each region.

```{r}
total_inventory <- c(
  sum(inventory_data$Region1),
  sum(inventory_data$Region2),
  sum(inventory_data$Region3)
)
total_inventory

```

# Calculate inventory surplus or shortage relative to demand.

```{r}
#If we have a small dataset
region1 <- sum(inventory_data$Region1) - demand_data2$Demand[1]
region1
region2 <- sum(inventory_data$Region2) - demand_data2$Demand[2]
region2
region3 <- sum(inventory_data$Region3) - demand_data2$Demand[3]
region3
#A large dataset we will use this R function for vector
surplus_shortage <- total_inventory - demand_data2$Demand
surplus_shortage

```

# Identify which region is most at risk of stock shortages.

```{r}
names(surplus_shortage) <- c("Region1", "Region2", "Region3")

surplus_shortage
names(surplus_shortage)[which.min(surplus_shortage)]
```

# Interpret the business implications.

Region 3 has the greatest risk of an inventory risk of shortage because its available inventory is lowest relative to its demand. This could result in stock outs and an inability to fulfill customer orders. The company should consider reallocating inventory from regions with larger surpluses to Region 3 or increasing future inventory shipments to Region 3.

# Part B: Shipment Matrix Analysis

## Using matrix operations:

1.  Create a shipment matrix from the data.

```{r}
#load shipments data
data3 <- "https://raw.githubusercontent.com/japhet125/Data_Science_605_Hws/refs/heads/main/shipments.csv"
shipment_data <- read_csv(data3)
shipment_data
shipment_matrix <- as.matrix(shipment_data[, 2:4])
shipment_matrix
```

1.  Compute row totals and column totals.

```{r}
#row total
warehouse_totals <- rowSums(shipment_matrix)
warehouse_totals
#column total
region_total <- colSums(shipment_matrix)
region_total
```

1.  Determine which warehouse ships the most products.

```{r}
names(warehouse_totals) <- c("A", "B", "C")
warehouse_totals
names(warehouse_totals)[which.max(warehouse_totals)]
```

1.  Determine which region receives the most products.

```{r}
names(region_total) <- c("Region1", "Region2", "Region3")
region_total
names(region_total)[which.max(region_total)]
```

## Visualization

```{r}
# Visualization 1: Inventory vs Demand by Region

region_summary <- data.frame(
  Region = c("Region1", "Region2", "Region3"),
  Inventory = total_inventory,
  Demand = demand_data2$Demand
)

region_summary_long <- region_summary %>%
  pivot_longer(
    cols = c(Inventory, Demand),
    names_to = "Type",
    values_to = "Units"
  )

ggplot(region_summary_long, aes(x = Region, y = Units, fill = Type)) +
  geom_col(position = "dodge") +
  labs(
    title = "Inventory vs Demand by Region",
    x = "Region",
    y = "Units",
    fill = "Category"
  ) +
  theme_minimal()
```

```{r}
# Visualization 2: Total Shipments by Warehouse

warehouse_summary <- data.frame(
  Warehouse = c("A", "B", "C"),
  Total_Shipments = warehouse_totals
)

ggplot(warehouse_summary,
       aes(x = Warehouse, y = Total_Shipments)) +
  geom_col() +
  labs(
    title = "Total Shipments by Warehouse",
    x = "Warehouse",
    y = "Total Products Shipped"
  ) +
  theme_minimal()
```

# Part C: Resource Allocation Model

Suppose inventory allocations must satisfy:

1.  **Express the system in matrix form**.

$$
\begin{bmatrix}
1 & 1 & 1 \\
2 & 1 & 0 \\
0 & 1 & 2
\end{bmatrix}
\begin{bmatrix}
x \\
y \\
z
\end{bmatrix}
=
\begin{bmatrix}
18 \\
26 \\
22
\end{bmatrix}
$$Rewrite the system as Ax=b :

$$
A = 
\begin{bmatrix}
1 & 1 & 1 \\
2 & 1 & 0 \\
0 & 1 & 2 
\end{bmatrix}
$$ $$
\mathbf{x} = 
\begin{bmatrix}
x \\
y \\
z
\end{bmatrix}
$$ $$
b = 
\begin{bmatrix}
18 \\
26 \\
22
\end{bmatrix}
$$

2\. **Solve the System Computationally**

```{r}
#creating the coef matrix
A <- matrix(c(
  1,1,1,
  2,1,0,
  0,1,2
  
), nrow = 3, byrow = TRUE)

A
#create b vector
b <- c(18,26,22)
b

#Solve for the unknown vector
#solution <- solve(A, b)
#solution
#det(A)
```

Solving x, y, z using R result to error due to the solution not having singular value but multiple values,

3.  **Verify the Solution**

    we have:

    x+y+z=18

    2x+y=26

    y+2z=22

    solution:

    x+y+z=18

    2x+y+0z=26

    0x+y+2z=22

    ------------------------------------------------------------------------

    3x+3y+3z=66

    ------------------------------------------------------------------------

    3(x+y+z)=66 =====\> (x+y+z)= 66/3 ====\> x+y+z=22

    x=22-y-z ====\> replace x by it value, 22-y-z+y+z= 18 ====\> 22=18 the system has no solution

    we could also try 2x+y+y+2z = 48 ====\> 2x+2y+2z = 48 then x+y+z = 24

    y =24 -x-z so if we replace the values we get x + 24 -x-z+z = 18 ==\> 24 = 18

    the system has no unique solution

recap: Since (24 \\neq 18), the equations are inconsistent. Therefore, there is no set of values for (x), (y), and (z) that satisfies all three equations.

This is also confirmed computationally in R because `solve(A, b)` returns a singular-system error and `det(A)` is equal to 0.

4.  **Business interpretation**

    Based on the analysis, the system of linear equations does not provide a valid number of truckloads that each warehouse should ship because the coefficient matrix is singular and the equations are inconsistent. Therefore, values for (x), (y), and (z) cannot be determined from the current system, and management would need to review the constraints before making a truckload allocation decision.

    However, the results from Parts A and B still provide useful information for management. In Part B, Warehouse A ships the most products with 1,200 units, while Warehouses B and C each ship 1,000 units. This shows that Warehouse A currently contributes the most to the shipment activity. Part A also shows that Region 3 has no inventory surplus, while Regions 1 and 2 each have a surplus of 300 units. Therefore, Region 3 is more at risk if demand increases.

    Mathematical models help management evaluate the needs of each region and the capacity of each warehouse so resources can be allocated efficiently. Systems of linear equations are useful in supply chain planning and logistics because they represent relationships between different constraints and help determine whether a proposed allocation is mathematically feasible.
