---
title: "Homework"
format: pdf_output
editor: visual
---

## Tasks

## Part 1 Inventory

```{r}
library(tidyverse)
#Inventory data load
data = ("https://raw.githubusercontent.com/japhet125/Data_Science_605_Hws/refs/heads/main/inventory.csv")
inventory_data = read.csv(data)
inventory_data
```

## Part 2 Demand

```{r}
#Demand data load
data2 = ("https://raw.githubusercontent.com/japhet125/Data_Science_605_Hws/refs/heads/main/demand.csv")
demand_data2 = read.csv(data2)
demand_data2
```

# Tasks

# Part A: Inventory Analysis

## Using vector operations:

```{r}
inventory_data$Region1
inventory_data$Region2
inventory_data$Region3
```

# Calculate total inventory available in each region.

```{r}
total_inventory <- c(
  sum(inventory_data$Region1),
  sum(inventory_data$Region2),
  sum(inventory_data$Region3)
)
total_inventory

```

# Calculate inventory surplus or shortage relative to demand.

```{r}
#If we have a small dataset
region1 <- sum(inventory_data$Region1) - inventory_data2$Demand[1]
region1
region2 <- sum(inventory_data$Region2) - inventory_data2$Demand[2]
region2
region3 <- sum(inventory_data$Region3) - inventory_data2$Demand[3]
region3
#A large dataset we will use this R function for vector
surplus_shortage <- total_inventory - inventory_data2$Demand
surplus_shortage

```

# Identify which region is most at risk of stock shortages.

```{r}
names(surplus_shortage) <- c("Region1", "Region2", "Region3")

surplus_shortage
names(surplus_shortage)[which.min(surplus_shortage)]
```

# Interpret the business implications.

Region 3 has the greatest risk of an inventory shortage because its available inventory is lowest relative to its demand. This could result in stockouts and an inability to fulfill customer orders. The company should consider reallocating inventory from regions with larger surpluses to Region 3 or increasing future inventory shipments to Region 3.

# Part B: Shipment Matrix Analysis

## Using matrix operations:

1.  Create a shipment matrix from the data.

```{r}
#load shipments data
data3 <- "https://raw.githubusercontent.com/japhet125/Data_Science_605_Hws/refs/heads/main/shipments.csv"
shipment_data <- read_csv(data3)
shipment_data
shipment_matrix <- as.matrix(shipment_data[, 2:4])
shipment_matrix
```

1.  Compute row totals and column totals.

```{r}
#row total
warehouse_totals <- rowSums(shipment_matrix)
warehouse_totals
#column total
region_total <- colSums(shipment_matrix)
region_total
```

1.  Determine which warehouse ships the most products.

```{r}
names(warehouse_totals) <- c("A", "B", "C")
warehouse_totals
names(warehouse_totals)[which.max(warehouse_totals)]
```

1.  Determine which region receives the most products.

```{r}
names(region_total) <- c("Region1", "Region2", "Region3")
region_total
names(region_total)[which.max(region_total)]
```

## Visualization

```{r}
# Visualization 1: Inventory vs Demand by Region

region_summary <- data.frame(
  Region = c("Region1", "Region2", "Region3"),
  Inventory = total_inventory,
  Demand = inventory_data2$Demand
)

region_summary_long <- region_summary %>%
  pivot_longer(
    cols = c(Inventory, Demand),
    names_to = "Type",
    values_to = "Units"
  )

ggplot(region_summary_long, aes(x = Region, y = Units, fill = Type)) +
  geom_col(position = "dodge") +
  labs(
    title = "Inventory vs Demand by Region",
    x = "Region",
    y = "Units",
    fill = "Category"
  ) +
  theme_minimal()
```

```{r}
# Visualization 2: Total Shipments by Warehouse

warehouse_summary <- data.frame(
  Warehouse = c("A", "B", "C"),
  Total_Shipments = warehouse_totals
)

ggplot(warehouse_summary,
       aes(x = Warehouse, y = Total_Shipments)) +
  geom_col() +
  labs(
    title = "Total Shipments by Warehouse",
    x = "Warehouse",
    y = "Total Products Shipped"
  ) +
  theme_minimal()
```

# Part C: Resource Allocation Model

Suppose inventory allocations must satisfy:

1.  Express the system in matrix form.

2.  Solve the system computationally.

3.  Verify the solution.

4.  Discuss how systems of equations arise in supply chain management.
